CF Registry Analysis

### 2022-12-23 Christine Chan/CF Registry/cf_registry_analysis.qmd
Update cf_registry_analysis.qmd

```{r}
# All data
df <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
df <- df %>%
  arrange(eDWID, ReviewYear) %>%
  mutate(`Age in 2018` = 2018 - (ReviewYear - Age_YrEnd))
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd <- df %>%
  filter(A_twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  slice_min(ReviewYear) %>%
  select(eDWID, ReviewYear, `Age in 2018`) %>%
  rename(first_cfrd = ReviewYear)
df <- left_join(df, first_cfrd, by = c("eDWID", "Age in 2018"))
df <- df %>%
  group_by(eDWID) %>%
  mutate(cfrd_ogtt = case_when(
    ReviewYear < first_cfrd ~ "No",
    ReviewYear >= first_cfrd ~ "Yes"
  )) %>%
  ungroup()
df$cfrd_ogtt[is.na(df$cfrd_ogtt)] <- "No"
df <- df %>%
  select(eDWID, ReviewYear, `Age in 2018`, first_cfrd, cfrd_ogtt, A_hgba1c)
ages <- df %>%
  group_by(eDWID) %>%
  summarise(`Age in 2018` = names(sort(table(`Age in 2018`), 
                                       decreasing = TRUE)[1]))
ages$`Age in 2018` <- round(as.numeric(ages$`Age in 2018`), 1)
```

```{r}
#| label: tbl-incidence
#| tbl-cap: "Year-on-Year Incidence of CFRD"
# Person-years variable
py <- 1000
# For each year, get the number of people diagnosed with CFRD that year
years <- 2003:2018
cases <- lapply(years, function(y) {
  total <- unique(first_cfrd$eDWID[first_cfrd$first_cfrd == y])
  cases_age_10 <- total[total %in% ages$eDWID[ages$`Age in 2018` < 10]]
  cases_age_10_18 <- total[total %in% ages$eDWID[ages$`Age in 2018` >= 10 &
    ages$`Age in 2018` < 19]]
  cases_age_19_30 <- total[total %in% ages$eDWID[ages$`Age in 2018` >= 19 &
    ages$`Age in 2018` < 31]]
  cases_age_30 <- total[total %in% ages$eDWID[ages$`Age in 2018` >= 31]]
  return(list(
    "total" = total, "cases_age_10" = cases_age_10,
    "cases_age_10_18" = cases_age_10_18,
    "cases_age_19_30" = cases_age_19_30,
    "cases_age_30" = cases_age_30
  ))
})
# Double check that there is no overlap between sets (there isn't!)
# lapply(1:length(cases), function(y){intersect(cases[[y]],unlist(cases[-y]))})
# For each year, get the number of people at risk
at_risk <- lapply(years, function(y) {
  ar <- unique(df$eDWID[df$ReviewYear == y])
  ar <- ar[!ar %in% unlist(cases[which(years <= y)]$total)]
  ar_age_10 <- intersect(
    df$eDWID[df$ReviewYear == y],
    ages$eDWID[ages$`Age in 2018` < 10]
  )
  ar_age_10_18 <- intersect(
    df$eDWID[df$ReviewYear == y],
    ages$eDWID[ages$`Age in 2018` >= 10 & ages$`Age in 2018` < 19]
  )
  ar_age_19_30 <- intersect(
    df$eDWID[df$ReviewYear == y],
    ages$eDWID[ages$`Age in 2018` >= 19 & ages$`Age in 2018` < 31]
  )
  ar_age_30 <- intersect(
    df$eDWID[df$ReviewYear == y],
    ages$eDWID[ages$`Age in 2018` >= 31]
  )
  return(list(
    "total" = ar, "ar_age_10" = ar_age_10, "ar_age_10_18" = ar_age_10_18,
    "ar_age_19_30" = ar_age_19_30, "ar_age_30" = ar_age_30
  ))
})
# Put everything in one table
inc_df <- lapply(years, function(y) {
  # Total
  new_cases <- length(cases[[which(years == y)]]$total)
  total_at_risk <- length(at_risk[[which(years == y)]]$total)
  total_inc <- new_cases / (total_at_risk / py)
  # < 10
  new_cases_10 <- length(cases[[which(years == y)]]$cases_age_10)
  at_risk_10 <- length(at_risk[[which(years == y)]]$ar_age_10)
  inc_10 <- new_cases_10 / (at_risk_10 / py)
  # 10 - 18
  new_cases_10_18 <- length(cases[[which(years == y)]]$cases_age_10_18)
  at_risk_10_18 <- length(at_risk[[which(years == y)]]$ar_age_10_18)
  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)
  # 19 - 30
  new_cases_19_30 <- length(cases[[which(years == y)]]$cases_age_19_30)
  at_risk_19_30 <- length(at_risk[[which(years == y)]]$ar_age_19_30)
  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)
  # > 30
  new_cases_30 <- length(cases[[which(years == y)]]$cases_age_30)
  at_risk_30 <- length(at_risk[[which(years == y)]]$ar_age_30)
  inc_30 <- new_cases_30 / (at_risk_30 / py)
  return(list(
    `Year` = y,
    `Total New Cases` = new_cases,
    `Total At Risk` = total_at_risk,
    `Total Incidence` = total_inc,
    `Total New Cases Age < 10` = new_cases_10,
    `Total At Risk Age < 10` = at_risk_10,
    `Total Incidence Age < 10` = inc_10,
    `Total New Cases Age 10-18` = new_cases_10_18,
    `Total At Risk Age 10-18` = at_risk_10_18,
    `Total Incidence Age 10-18` = inc_10_18,
    `Total New Cases Age 19-30` = new_cases_19_30,
    `Total At Risk Age 19-30` = at_risk_19_30,
    `Total Incidence Age 19-30` = inc_19_30,
    `Total New Cases Age > 30` = new_cases_30,
    `Total At Risk Age > 30` = at_risk_30,
    `Total Incidence Age > 30` = inc_30
  ))
})
inc_df <- as.data.frame(do.call(rbind, inc_df))
inc_df <- data.frame(lapply(inc_df, as.numeric), check.names = F)
kable(inc_df, digits = 3)
```

@@

```{r}
df$cfrd_ogtt <- factor(df$cfrd_ogtt, levels = c("No", "Yes"), ordered = T)
cfrd <- df %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ever = max(cfrd_ogtt),
    time_from_diag = ReviewYear - first_cfrd
  ) %>%
  filter(cfrd_ever == "Yes", time_from_diag <= 0, time_from_diag >= -1) %>%
  summarise(hba1c = last(na.omit(round(A_hgba1c, 1)))) %>%
  drop_na()
non_cfrd <- df %>%
  group_by(eDWID) %>%
  mutate(cfrd_ever = max(cfrd_ogtt)) %>%
  filter(cfrd_ever == "No") %>%
  summarise(hba1c = last(na.omit(round(A_hgba1c, 1)))) %>%
  drop_na()
```

@@

>>>>

```{r}
# All data
encounter <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_encountersMerged_Del1.csv",
  na.strings = c("", "N/A")
)
annual <- read.csv("Z:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Laura/Peds Endo/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
encounter <- encounter %>%
  arrange(eDWID, reviewyear, encounternum)
# Add age
ages <- annual %>%
  group_by(eDWID) %>%
  summarise(`Age in 2018` = names(sort(table(`Age in 2018`), 
                                       decreasing = TRUE)[1]))
encounter = left_join(encounter,ages,by = "eDWID")
# For those with CFRD by OGTT, find first visit with diagnosis
first_cfrd <- encounter %>%
  filter(twohour_bloodglucose >= 200) %>%
  group_by(eDWID) %>%
  slice_min(reviewyear) %>%
  select(eDWID, reviewyear, `Age in 2018`) %>%
  rename(first_cfrd = reviewyear)
encounter <- left_join(encounter, first_cfrd, by = c("eDWID", "Age in 2018"))
encounter <- encounter %>%
  group_by(eDWID) %>%
  mutate(cfrd_ogtt = case_when(
    reviewyear < first_cfrd ~ "No",
    reviewyear >= first_cfrd ~ "Yes"
  )) %>%
  ungroup()
encounter$cfrd_ogtt[is.na(df$cfrd_ogtt)] <- "No"
```

```{r}
#| label: tbl-incidence
#| tbl-cap: "Year-on-Year Incidence of CFRD"
# Person-years variable
py <- 1000
# For each year, get the number of people diagnosed with CFRD that year
years <- 2003:2018
cases <- lapply(years, function(y) {
  total <- unique(first_cfrd$eDWID[first_cfrd$first_cfrd == y])
  cases_age_10 <- total[total %in% ages$eDWID[ages$`Age in 2018` < 10]]
  cases_age_10_18 <- total[total %in% ages$eDWID[ages$`Age in 2018` >= 10 &
    ages$`Age in 2018` < 19]]
  cases_age_19_30 <- total[total %in% ages$eDWID[ages$`Age in 2018` >= 19 &
    ages$`Age in 2018` < 31]]
  cases_age_30 <- total[total %in% ages$eDWID[ages$`Age in 2018` >= 31]]
  return(list(
    "total" = total, "cases_age_10" = cases_age_10,
    "cases_age_10_18" = cases_age_10_18,
    "cases_age_19_30" = cases_age_19_30,
    "cases_age_30" = cases_age_30
  ))
})
# Double check that there is no overlap between sets (there isn't!)
# lapply(1:length(cases), function(y){intersect(cases[[y]],unlist(cases[-y]))})
# For each year, get the number of people at risk
at_risk <- lapply(years, function(y) {
  ar <- unique(encounter$eDWID[encounter$reviewyear == y])
  ar <- ar[!ar %in% unlist(cases[which(years <= y)]$total)]
  ar_age_10 <- intersect(
    encounter$eDWID[encounter$reviewyear == y],
    ages$eDWID[ages$`Age in 2018` < 10]
  )
  ar_age_10_18 <- intersect(
    encounter$eDWID[encounter$reviewyear == y],
    ages$eDWID[ages$`Age in 2018` >= 10 & ages$`Age in 2018` < 19]
  )
  ar_age_19_30 <- intersect(
    encounter$eDWID[encounter$reviewyear == y],
    ages$eDWID[ages$`Age in 2018` >= 19 & ages$`Age in 2018` < 31]
  )
  ar_age_30 <- intersect(
    encounter$eDWID[encounter$reviewyear == y],
    ages$eDWID[ages$`Age in 2018` >= 31]
  )
  return(list(
    "total" = ar, "ar_age_10" = ar_age_10, "ar_age_10_18" = ar_age_10_18,
    "ar_age_19_30" = ar_age_19_30, "ar_age_30" = ar_age_30
  ))
})
# Put everything in one table
inc_df <- lapply(years, function(y) {
  # Total
  new_cases <- length(cases[[which(years == y)]]$total)
  total_at_risk <- length(at_risk[[which(years == y)]]$total)
  total_inc <- new_cases / (total_at_risk / py)
  # < 10
  new_cases_10 <- length(cases[[which(years == y)]]$cases_age_10)
  at_risk_10 <- length(at_risk[[which(years == y)]]$ar_age_10)
  inc_10 <- new_cases_10 / (at_risk_10 / py)
  # 10 - 18
  new_cases_10_18 <- length(cases[[which(years == y)]]$cases_age_10_18)
  at_risk_10_18 <- length(at_risk[[which(years == y)]]$ar_age_10_18)
  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)
  # 19 - 30
  new_cases_19_30 <- length(cases[[which(years == y)]]$cases_age_19_30)
  at_risk_19_30 <- length(at_risk[[which(years == y)]]$ar_age_19_30)
  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)
  # > 30
  new_cases_30 <- length(cases[[which(years == y)]]$cases_age_30)
  at_risk_30 <- length(at_risk[[which(years == y)]]$ar_age_30)
  inc_30 <- new_cases_30 / (at_risk_30 / py)
  return(list(
    `Year` = y,
    `Total New Cases` = new_cases,
    `Total At Risk` = total_at_risk,
    `Total Incidence` = total_inc,
    `Total New Cases Age < 10` = new_cases_10,
    `Total At Risk Age < 10` = at_risk_10,
    `Total Incidence Age < 10` = inc_10,
    `Total New Cases Age 10-18` = new_cases_10_18,
    `Total At Risk Age 10-18` = at_risk_10_18,
    `Total Incidence Age 10-18` = inc_10_18,
    `Total New Cases Age 19-30` = new_cases_19_30,
    `Total At Risk Age 19-30` = at_risk_19_30,
    `Total Incidence Age 19-30` = inc_19_30,
    `Total New Cases Age > 30` = new_cases_30,
    `Total At Risk Age > 30` = at_risk_30,
    `Total Incidence Age > 30` = inc_30
  ))
})
inc_df <- as.data.frame(do.call(rbind, inc_df))
inc_df <- data.frame(lapply(inc_df, as.numeric), check.names = F)
kable(inc_df, digits = 3)
```

@@

```{r}
#| label: tbl-prevalence
#| tbl-cap: "Yearly Prevalence of CFRD"
df %>% group_by(ReviewYear) %>% 
  summarise(Prevalance = mean(cfrd_status_annual==3,na.rm = T))

t = df %>% select(eDWID,ReviewYear,A_twohour_bloodglucose,cfrd_ogtt,cfrd_status_annual)
```

In @tbl-prevalence, prevalence is calculated as the total number of CFRD cases seen in clinic divided by the total number of people seen in clinic in a given year.

```{r}
# Prepare data for ROC
df$cfrd_ogtt <- factor(df$cfrd_ogtt, levels = c("No", "Yes"), ordered = T)
cfrd <- df %>%
  group_by(eDWID) %>%
  mutate(
    cfrd_ever = max(cfrd_ogtt),
    time_from_diag = ReviewYear - first_cfrd
  ) %>%
  filter(cfrd_ever == "Yes", time_from_diag <= 0, time_from_diag >= -1) %>%
  summarise(hba1c = last(na.omit(round(A_hgba1c, 1)))) %>%
  drop_na()
non_cfrd <- df %>%
  group_by(eDWID) %>%
  mutate(cfrd_ever = max(cfrd_ogtt)) %>%
  filter(cfrd_ever == "No") %>%
  summarise(hba1c = last(na.omit(round(A_hgba1c, 1)))) %>%
  drop_na()
```

@@

# Questions

1.  For these calculations, do we need to just use those with an OGTT? Or everybody in the database?
